---
title: "Greenhouse Gas Emissions 2018"
author: "Gabriel Berlin"
date: "11/29/2019"
output: html_notebook
---

## load libraries
```{r}
# clear environment
rm(list = ls()) 

# load pakcages
library(DataComputing)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(mosaic)
library(mosaicData)
library(readr)
```

## load data
```{r}
fullEmissionsTable <- read_csv("/home/gabriel/college/Greenhouse Gas Emissions 2018/direct_emitters.csv", skip = 3)
```

## view data
```{r}
names(fullEmissionsTable)
str(fullEmissionsTable)
head(fullEmissionsTable, 10)
```

The table's columns contain information about three different things: information about the facility, actual emissions data, and how gasses were emitted.
How gasses were emitted is not relevant to the research question, so I will focus on only the first two categories and I will get rid of the variables I do not need..

# data wrangling
```{r}
emissionsTable <- 
  fullEmissionsTable %>%
  rename(sector = `Industry Type (sectors)`) %>%
  select(`Facility Id`, `FRS Id`, `Facility Name`, `State`, Latitude, Longitude, 
         `sector`, `Total reported direct emissions`, 
         `CO2 emissions (non-biogenic)`, `Methane (CH4) emissions`, 
         `Nitrous Oxide (N2O) emissions`, `HFC emissions`, `PFC emissions`, `SF6 emissions`,
         `NF3 emissions`, `Other Fully Fluorinated GHG emissions`, `HFE emissions`,
         `Other GHGs (metric tons CO2e)`, `Biogenic CO2 emissions (metric tons)`)
```


# create list of sectors/industries
```{r}
sector <-
  emissionsTable %>%
  group_by(sector) %>%
  summarise(total = n()) %>%
  select(sector, total)
sector
```

There are multiple sectors in some of the values. For example, "Chemicals,Coal-based Liquid Fuel Supply,Suppliers of CO2". Because there are multiple values in a single cell, this data is not in tidy form. To make it in tidy form I will make sure there is only one value in a cell.
I will need to use regular expressions and a loop to replace the values.

### pseudo-code: first create a regular expression, then use the loop
```{r}
chem <- "Chemicals"
metals <- "Metals"
minerals <- "Minerals"
natural <- "Natural Gas"
other <- "Other"
petroSys <- "Petroleum and "
petroSup <- "Petroleum Product"
paper <- "Pulp"
injection <- "Injection of "

for (s in sector) {
  if (s == "$Chemical") {
    mutate(s == chem)
  } else if (s == )
}
```


